The Design and Implementation of the Synchronous Language \CEU

### 2016-04-02 paper.tex
...

The \code{await} statement of \CEU supports wall-clock time and handles deltas 
automatically, resulting in more robust applications.
In the example in Figure~\ref{lst.timers}.a, suppose that after the first 
\code{await} request, the underlying system gets busy and takes 15ms to check 
for expiring awaits.
The \CEU scheduler will notice that the \code{await 10ms} has not only already 
expired, but is delayed with \code{delta=5ms}.
Then, the awaiting trail awakes, sets \code{v=1}, and invokes \code{await 1ms}.
As the current delta is still higher than the requested timeout (i.e. $5ms > 
1ms$), the trail is rescheduled for execution, now with \code{delta=4ms}.

\begin{figure}
\begin{minipage}[t]{0.45\linewidth}
\begin{lstlisting}
var int v;
await 10ms;
v = 1;
await 1ms;
v = 2;

.
\end{lstlisting}
\centering\small{\ax}
\end{minipage}
%
\begin{minipage}[t]{0.51\linewidth}
\begin{lstlisting}
par/or do
    await 10ms;
    <...>         // any non-awaiting sequence
    await  1ms;
    v = 1;
with
    await 12ms;
    v = 2;
end
\end{lstlisting}
\centering\small{\bx}
\end{minipage}
%\rule{8.5cm}{0.37pt}
\caption{ First-class timers in \CEU.
\label{lst.timers}
}
\end{figure}

\CEU also considers that time is a physical quantity that can be added and 
compared.
For instance, for the example in Figure~\ref{lst.timers}.b, although the 
scheduler cannot guarantee that the first trail terminates exactly in 11ms, it 
can at least ensure that the program always terminates with \code{v=1}.
%
Given that any non-awaiting sequence is considered to take no time in the 
synchronous model, the first trail is guaranteed to terminate before the second 
trail, because $10+1 < 12$.
%
A similar program in a language without first-class support for timers would 
depend on the execution timings for the code marked as \code{<...>}, making the 
reasoning about the execution behavior more difficult.

>>>>

The \code{await} statement of \CEU supports wall-clock time and handles deltas 
automatically, resulting in more robust applications.
In the example in Figure~\ref{lst.timers}.a, suppose that after the first 
\code{await} request, the underlying system gets busy and takes 15ms to check 
for expiring awaits.
The \CEU scheduler will notice that the \code{await 10ms} (ln. 2) has not only 
already expired, but is delayed with \code{delta=5ms}.
Then, the awaiting trail awakes, sets \code{v=1} (ln. 3), and invokes 
\code{await 1ms} (ln. 4).
As the current delta is still higher than the requested timeout (i.e. $5ms > 
1ms$), the trail is rescheduled for execution, now with \code{delta=4ms}.

\begin{figure}
\begin{minipage}[t]{0.45\linewidth}
\begin{lstlisting}[numbers=left,xleftmargin=3.5em]
var int v;
await 10ms;
v = 1;
await 1ms;
v = 2;

.
\end{lstlisting}
\centering\small{\ax}
\end{minipage}
%
\begin{minipage}[t]{0.55\linewidth}
\begin{lstlisting}[numbers=left,xleftmargin=3.5em]
par/or do
    await 10ms;
    <...>       // any non-awaiting sequence
    await  1ms;
    v = 1;
with
    await 12ms;
    v = 2;
end
\end{lstlisting}
\centering\small{\bx}
\end{minipage}
%\rule{8.5cm}{0.37pt}
\caption{ First-class timers in \CEU.
\label{lst.timers}
}
\end{figure}

\CEU also considers that time is a physical quantity that can be added and 
compared.
For instance, for the example in Figure~\ref{lst.timers}.b, although the 
scheduler cannot guarantee that the first trail terminates exactly in 11ms (ln.  
2,4), it can at least ensure that the program always terminates with 
\code{v=1}.
%
Given that any non-awaiting sequence is considered to take no time in the 
synchronous model, the first trail (ln. 2--5) is guaranteed to terminate before 
the second trail (ln. 7--8), because $10+1 < 12$.
%
A similar program in a language without first-class support for timers would 
depend on the execution timings for the code marked as \code{<...>}, making the 
reasoning about the execution behavior more difficult.
